A Large and Variable Leading Tail of Helium in a Sub-Saturn Undergoing Runaway Inflation

### 2023-07-07 paper/paper1/ms.tex
comments from Suvrath and ZJ

The Habitable Zone Planet Finder Spectrograph \citep[HPF;][]{2012SPIE.8446E..1SM,2014SPIE.9147E..1GM, 2019Optic...6..233M} on the queue-scheduled 10-meter \emph{Hobby-Eberly Telescope} \citep[HET;][]{1998SPIE.3352...34R} operates in the near-IR from $8100-12800~$\AA~ spanning the \textit{z}, \textit{Y}, and \textit{J} bands at spectral resolving power $R=55,000$. The HET fixed-elevation design \citep{2007PASP..119..556S} limits the observability of \object{HAT-P-67} to less than 1 hour ``tracks'' for a fixed range of hour angles before (east track) and after (west track) the star transits the meridian.  Whereas conventional steerable telescopes could conduct continuous point-and-stare observations of \object{HAT-P-67} for hours, HET cannot.  In practice, this limitation means that in-transit and out-of-transit observational phases were rarely possible on the same night.  Instead, we organized the observations into four campaigns to coincide with \object{HAT-P-67 b} transits on the nights beginning 2020 April 27, 2020; May 21, 2020; June 14, 2020; and 2022 April 28.  These campaigns have out-of-transit observations at least one night before and one night after, and often two nights on either side of the transit.  Two more transit snapshots were obtained in 2022 June-July without the accompanying visits immediately before and after.  The in-transit campaigns had up to 14 exposures per HET track, with integration times between 5-8.5 minutes.  We also obtained random-in-phase lower-priority P1-4 reconnaissance observations \citep{2007PASP..119..556S}.  These out-of-transit snapshot observations typically received 4 or fewer individual exposures.  We observed HAT-P-67 with HPF for a total of 41 visits on 39 unique nights, with two of those nights observing both the east and west HET tracks.  The total on-source integration time exceeds 13.8 hours, over $N=152$ individual spectra.  Table \ref{tabHPFLog} summarizes the log of observations.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT_P_67b_He_spectrum.png}
    \caption{Overlay of all 152 individual HPF exposures of HAT-P-67, spanning 2020-2022. Variability is seen in the \ion{He}{1} 10833 \AA~ triplet, denoted by the vertical orange shaded band, but not in the adjacent regions.  These snapshot spectra were barycentric corrected, and continuum flattened to the regions in the blue vertical bands.  Sharp telluric absorption lines have been masked in regions near 10835 and 10837.5 \AA.}
    \label{fig:HPFheliumOverview}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT_P_67b_Helium10830EW_byCampaign.png}
    \caption{Equivalent Width lightcurve of Helium absorption in HAT-P-67b.  The system exhibits a characteristic absorption leading up to transit, followed by a sharp decline after transit passage.}
    \label{fig:HPFtimeseries}
\end{figure}

>>>>

The Habitable Zone Planet Finder Spectrograph \citep[HPF;][]{2012SPIE.8446E..1SM,2014SPIE.9147E..1GM, 2019Optic...6..233M} on the queue-scheduled 10-meter \emph{Hobby-Eberly Telescope} \citep[HET;][]{1998SPIE.3352...34R} operates in the near-IR from $8100-12800~$\AA~ spanning the \textit{z}, \textit{Y}, and \textit{J} bands at spectral resolving power $R=55,000$. The HET fixed-elevation design \citep{2007PASP..119..556S} limits the observability of \object{HAT-P-67} to less than 1 hour ``tracks'' for a fixed range of hour angles before (east track) and after (west track) the star transits the meridian.  Whereas conventional steerable telescopes could conduct continuous point-and-stare observations of \object{HAT-P-67} for hours, HET cannot.  In practice, this limitation means that in-transit and out-of-transit observational phases were rarely possible on the same night.  Instead, we organized the observations into four campaigns to coincide with \object{HAT-P-67 b} transits on the nights beginning 2020 April 27, 2020; May 21, 2020; June 14, 2020; and 2022 April 28.  These campaigns have out-of-transit observations at least one night before and one night after, and often two nights on either side of the transit.  Two more transit snapshots were obtained in 2022 June-July without the accompanying visits immediately before and after.  The in-transit campaigns had up to 14 exposures per HET track, with integration times between 5-8.5 minutes.  We also obtained random-in-phase lower-priority P1-4 reconnaissance observations \citep{2007PASP..119..556S}.  These out-of-transit snapshot observations typically received 4 or fewer individual exposures.  We observed HAT-P-67 with HPF for a total of 41 visits on 39 unique nights, with two of those nights observing both the east and west HET tracks.  The total on-source integration time exceeds 13.8 hours, with $N=152$ individual spectra possessing typical signal-to-noise ratio of 80 per pixel.  Table \ref{tabHPFLog} summarizes the log of observations.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT_P_67b_He_spectrum.png}
    \caption{Overlay of all 152 individual HPF exposures of HAT-P-67, spanning 2020-2022. Variability is seen in the \ion{He}{1} 10833 \AA~ triplet near the vertical orange shaded band, but not in the adjacent Si line.  These snapshot spectra were barycentric corrected, and continuum flattened with a linear fit to the regions in the blue vertical bands.  Sharp telluric absorption lines have been masked in regions near 10835 and 10837.5 \AA.}
    \label{fig:HPFheliumOverview}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/HAT_P_67b_Helium10830EW_byCampaign.png}
    \caption{Equivalent Width lightcurve of Helium absorption in HAT-P-67b.  The system exhibits a characteristic absorption leading up to transit, followed by a sharp decline after transit passage.  The EWs were computed in the orange shaded band (10832.3$-$10834.2 \AA) in Figure \ref{fig:HPFheliumOverview}.}
    \label{fig:HPFtimeseries}
\end{figure}
